Think deep for reef resilience

### 2025-08-15 paper/paper_draft.qmd
Combine paras
Paragraphs are three sentences minimum
Move carbonate budget results
Flow should match the order in which they were introduced in the method section

ADRIA-CoralBlox was run for 4096 scenarios for each of the five GCMs (20,480 total model
runs). Here, a scenario represents a unique combination of assumed environmental (namely one
of the five DHW trajectories) and ecological factors including coral growth and mortality
rates, coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol'
sampling method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a random-Quasi
Monte Carlo approach which balances scenario exploration with improved convergence
properties compared to pseudo-random sampling. A detailed comparative analysis of the
sampling methods can be found in [@kucherenko2015].

For the purpose of scenario assessment, the resulting timeseries for each model run were
scaled by the total area of the reef to represent cover as a percentage of the total
estimated reef area. These timeseries of coral cover were then clustered based on their
temporal behaviour.

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2075. The assessed carbonate budget thresholds range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/depth_carbonate_scatter.png){#fig-access-esm1-5-depth-carbonate
fig-align="center" width="14.82cm"}

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW timeseries
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor). Conversely, there are no consistent patterns across
bioregions in reef area-relative cover and mean timeseries DHW levels, with clusters having
similar mean DHW levels in many bioregions (@sfig-access-esm1-5-bioregion-dhwmean).
Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels @sfig-access-esm1-5-bioregion-dhwtol). There are no consistent differences
across bioregions and GCMs between clusters in their reefs' initial proportion of habitable
area occupied by corals (@sfig-access-esm1-5-bioregion-initialproportion). Across most
bioregions there are not consistent differences between clusters in connectivity metrics
such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi) and total
connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).

>>>>

ADRIA-CoralBlox was run for 4096 scenarios for each of the five GCMs (20,480 total model
runs). Here, a scenario represents a unique combination of assumed environmental (namely one
of the five DHW trajectories) and ecological factors including coral growth and mortality
rates, coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol'
sampling method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a random-Quasi
Monte Carlo approach which balances scenario exploration with improved convergence
properties compared to pseudo-random sampling. A detailed comparative analysis of the
sampling methods can be found in [@kucherenko2015]. For the purpose of scenario assessment,
the resulting timeseries for each model run were scaled by the total area of the reef to
represent cover as a percentage of the total estimated reef area. These timeseries of coral
cover were then clustered based on their temporal behaviour.

@@

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW timeseries
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor). Conversely, there are no consistent patterns across
bioregions in reef area-relative cover and mean timeseries DHW levels, with clusters having
similar mean DHW levels in many bioregions (@sfig-access-esm1-5-bioregion-dhwmean).
Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels (@sfig-access-esm1-5-bioregion-dhwtol). There are no consistent differences
across bioregions and GCMs between clusters in their reefs' initial proportion of habitable
area occupied by corals (@sfig-access-esm1-5-bioregion-initialproportion). Across most
bioregions there are not consistent differences between clusters in connectivity metrics
such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi) and total
connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2075. The assessed carbonate budget thresholds range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/depth_carbonate_scatter.png){#fig-access-esm1-5-depth-carbonate
alt="Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to 2075. The assessed carbonate budget thresholds range from 10 to 20% of total reef area. Bracketed x-axis values represent the Spearman's correlation between reef depth and the number of years cover exceeds the carbonate budget threshold. Point colour represents the median reef depth level in metres. Displays results from median timeseries using ACCESS-ESM1-5."
fig-align="center" width="14.82cm"}

## Partial Dependence

The bioregion variable was excluded from partial dependence analysis due to its categorical
nature providing limited interpretable information. The decision to exclude bioregions from
the analysis is also justified by a low feature importance score (discussed
